---
title: "[Project]Taichung_Gastronomy_Map"
author: "Dingyi Lai"
date: "2018/1/3"
output: html_document
---

# Introduction
As a special municipality being famed for its delicious food, Taichung attracts a large number of tourists every year. But where could we accommodate to own the most practical traffic for gastronomy?

# Goal
Find the best accommodation point for gastronomy in Taichung.

# Procedure
## Clean the platform
```{r}
rm(list = ls()) 
```

## Change the directory
```{r, eval=FALSE}
# setwd("/Users/aubrey/Documents/GitHub/Data-Science/")
```

## Read the data
```{r}

data <- read.csv("[FCU]R/清洗資料經緯度new.csv",header = TRUE,stringsAsFactors = FALSE,fileEncoding = "BIG5")
```

## Change columns name
```{r}
colnames(data)[4:5] <- c("lat","lon")
```

## Observe the structure of data
```{r}
class(data)
```

## Install necessary packages
```{r,message=FALSE}
library(dplyr)
library(magrittr)
library(factoextra)
library(rjson)
library(ggmap)
library(ggplot2)
library(sp)
```

## Clean data, limit the latitude and longitude and store it as new
```{r}
data%>%
  filter(lon>120.58 & lon<120.697 & lat<24.1901 & lat>24.133)->new
```

## kmeans clustering
```{r}
km <- kmeans(new[,c(4,5)],centers = 4)
```

## Store the results into new
```{r}
new$cluster <- km$cluster
```

## Observe result of clustering
```{r}
table(new$cluster,new$分類)
```

## Clustering Visualization
```{r,warning=FALSE}
fviz_cluster(km,data =new[,c(4,5)],geom = c("point","text"), 
             ellipse.type = "norm")
```

## Find the median of latitude and longitude for each group
```{r}
lat.med <- c()
lon.med <- c()
for(i in 1:4){
  
  lat.med[i] <- median(new[new$cluster==i,]$lat)
  lon.med[i] <- median(new[new$cluster==i,]$lon)
}
```

## Store them into data.frame as knew
```{r}
knew <- data.frame(lat=lat.med,lon=lon.med)
```

## Point Plot for grouped data
```{r,warning=FALSE}
qmap(location =c(lat=median(new$lat),lon=median(new$lon)) ,zoom = 12,language ="zh-TW")+
  geom_point(aes(x=lon,y=lat,col=new$cluster,size=1,alpha=0.6),data = new)+
  theme(legend.position = "none")+
  ggtitle("Scatter plot of the classified addresses")
```

## Point Plot for clustering
```{r,warning=FALSE}
qmap(location =c(lat=median(knew$lat),lon=median(knew$lon)) ,zoom = 12,language ="zh-TW")+
  geom_point(aes(x=lon,y=lat,col=rainbow(nrow(knew)),alpha=0.6),size=5,data = knew)+
  theme(legend.position = "none")+
  ggtitle("Scatter plot of the kmeans result")
```

## Zoom in
```{r,warning=FALSE}
qmap(location =c(lat=median(knew$lat),lon=median(knew$lon)) ,zoom = 12,language ="zh-TW")+
  geom_point(aes(x=lon,y=lat,col=rainbow(nrow(knew)),alpha=0.6),size=70,data = knew)+
  theme(legend.position = "none")+
  ggtitle("Scatter plot of the kmeans result")
```
