---
title: "Multivariate Gauss and Testing"
author: "Dingyi Lai"
date: "11/28/2020"
output:
  html_document:
    includes:
    in_header: header.tex
    latex_engine: xelatex
    toc: true
    depth: 3
    number_sections: true
    theme: united
    highlight: tango
    toc_float: true
    fontsize: "12pt"
    papersize: "a5"
    geometry: "margin=1in"
---

#   multivariate Gauss

```{r}
## function to plot 2D N
nv2.3d <- function(sigma1,sigma2,rho, phi, theta){
  
  library(mvtnorm)
  
  #Set MU and COV
  mu <- rep(0,2)
  Sigma <- matrix(c(sigma1^2,rho*sigma1*sigma2,rho*sigma1*sigma2,sigma2^2),ncol=2,nrow=2)
  
  #Set grid to compute density
  x1 <- x2 <- seq(-5,5,0.25)
  grid <- expand.grid(x1,x2)
  
  fvals <- matrix(dmvnorm(grid,mean=mu,sigma=Sigma),ncol=length(x1),nrow=length(x2))
  
  #3D plot of density
  persp(x=x1,y=x2,z=fvals,ticktype="detailed",xlab="x_1",ylab="x_2",phi=phi,theta=theta)
  title(main=paste("sigma1=",as.character(sigma1), ", sigma2=",as.character(sigma2), ", rho=",as.character(rho)))
}

## function to compute countours
nv2.contour <- function(sigma1,sigma2,rho){
  
  library(mvtnorm)
  
  #Set MU and COV
  mu <- rep(0,2)
  Sigma <- matrix(c(sigma1^2,rho*sigma1*sigma2,rho*sigma1*sigma2,sigma2^2),ncol=2,nrow=2)
  
  #Set grid
  x1 <- x2 <- seq(-5,5,0.25)
  grid <- expand.grid(x1,x2)
  
  fvals <- matrix(dmvnorm(grid,mean=mu,sigma=Sigma),ncol=length(x1),nrow=length(x2))
  
  #plot heights
  contour(x=x1,y=x2,z=fvals,xlab="x_1",ylab="x_2")
  title(main=paste("sigma1=",as.character(sigma1), ", sigma2=",as.character(sigma2), ", rho=",as.character(rho)))
}

```


##    Manipulate version
```{r,eval=FALSE}
    
## Use slides for parameters
library(manipulate)

########
par(mfrow=c(1,1))
manipulate(nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta), rho = slider(-.9,.9,step=0.1), phi = slider(0,90,step=10), theta = slider(0,90,step=10))

manipulate(nv2.contour(sigma1,sigma2, rho), rho = slider(-.9,.9,step=0.1),sigma1 =slider(0.5,4,step=0.5),
           sigma2 = slider(0.5,4,step=0.5))
           

```

##    Normal version

```{r}

rho = -.9
phi = 0
theta = 0
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

rho = -.9
phi = 90
theta = 90
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

nv2.contour(sigma1 = 1,sigma2 = 3, rho)

rho = -.5
phi = 0
theta = 0
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

rho = -.5
phi = 90
theta = 90
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

nv2.contour(sigma1 = 1,sigma2 = 3, rho)

rho = 0
phi = 0
theta = 0
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

rho = 0
phi = 0
theta = 90
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

rho = 0
phi = 90
theta = 0
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

rho = 0
phi = 90
theta = 90
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

nv2.contour(sigma1 = 1,sigma2 = 3, rho)

rho = 0.5
phi = 0
theta = 0
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

nv2.contour(sigma1 = 1,sigma2 = 3, rho)

rho = 0.9
phi = 0
theta = 0
nv2.3d(sigma1 = 1,sigma2 = 3, rho, phi, theta)

nv2.contour(sigma1 = 1,sigma2 = 3, rho)

```



#   another contour

recall formula multiv. Gaussian

```{r}
library(mvtnorm)
set.seed(1234)
s1 <- rmvnorm(1000, c(2, 3), matrix(c(10^2, 6^2, 6^2, 6^2),2,2))
mu <- apply(s1, 2, mean)   ## sample mean vector
sigma <- cov(s1)               ## sample variance-covariance matrix  


sigma.inv = solve(sigma, matrix(c(1,0,0,1),2,2))

ellipse <- function(s,t){
                    u<-c(s,t)-mu; 
                    e <- (-1) * (u %*% sigma.inv %*% u) / 2; 
                    exp(e)/(2 * pi * sqrt(det(sigma.inv)))}         
                    
                    
n <- 60
x <- (0:(n-1))*2 - 50
y <- (0:(n-1))*2 - 50

z <- mapply(ellipse, as.vector(rep(x,n)), as.vector(outer(rep(0,n), y, `+`)))

plot(s1, pch=21, xlim=c(-35, 35), ylim=c(-35,35), xlab="Margin X 1", ylab="Margin X 2", bg = "lightgray", cex = 0.3)
contour(x,y,matrix(z,n,n), levels=(0:15), col = terrain.colors(16), add=TRUE, lwd = 1)

```

#   conditional distribution

```{r}

Sigma <- matrix(c(1,.8,.8,1), nrow=2) ## variance-covariance matrix

x <- seq(-3,3,0.01)
contour(x,x,outer(x,x,function(x,y){dmvnorm(cbind(x,y),sigma=Sigma)}), col = "blue")

abline(v=.7, lwd=2, lty=2, col = "red")
text(0.75, -2, labels=expression(x[1]==0.7), col = "red", pos = 4)

### conditional distribution of X2 | X1 = 0.7
y <- dnorm(x, mean =  0.8 * 0.7, sd = sqrt(1 - 0.8^2))
lines(y-abs(min(x)),x,lty=2,lwd=2, col = "red")

### marginals
m1 <- m2 <- dnorm(x, 0, 1)
lines(x, m1 - abs(min(x)), lty = 1, lwd = 2, col = "gray30")
lines(m2 - abs(min(x)), x, lty = 1, lwd = 2, col = "gray30")


## (X_2|X_1 = -1)
contour(x,x,outer(x,x,function(x,y){dmvnorm(cbind(x,y),sigma=Sigma)}), col = "blue")
abline(v=-1, lwd=2, lty=2, col = "red")

### conditional distribution of X2 | X1 = - 1
y2 <- dnorm(x, mean = 0.8 * (- 1), sd = sqrt(1 - 0.8^2))
lines(-y2 + max(x),x,lty=2,lwd=2, col = "red")

#### same thing for set of different x (conditional distributions)
contour(x,x,outer(x,x,function(x,y){dmvnorm(cbind(x,y),sigma=Sigma)}), col= "blue")
condN <- function(x, cx){dnorm(x, mean = 0.8 * cx, sd = sqrt(1 - 0.8^2))}
for (i in seq(-2, 2, by = 0.25)){col <- colors()[grep("red",colors())][4*i + 9]; 
                                 lines(condN(x, i) + i, x, lwd = 2, col = col);
                                 abline(v = i, lwd=1, lty=1, col = col)}
```


#    wishart

```{r}
# W_1
S <- as.matrix(1)
sample <- rWishart(10, df = 10, S)     

### W Vs. Chi^2
set.seed(1234)
sampleWishart <- rWishart(5000, df = 10, S)
sampleChiSq <- rchisq(5000, df = 10)

par(mfrow = c(1,2))
hist(sampleWishart, col = "lightblue", main = expression(paste("Wishart Distribution",
               sep ="")), xlim = c(0, 40), breaks = 30, freq = F)
lines(density(sampleWishart), col = "red", lwd = 2)
hist(sampleChiSq, col = "lightblue", main = expression(paste(chi^2, "Distribution",
 sep = "")), xlim = c(0,40), breaks = 30, freq = F)
lines(density(sampleChiSq), col = "red", lwd = 2)
```

#   manova

##    data1
```{r}
## Hotelling's $T^2$ test 
Tibet <- source("tibetskulls.txt")$value
Tibet

# Using Hotelling's $T^2$ test, we test the hypothesis that the five-dimensional mean vectors of skull
# measurements are the same in the two populations from which the sample arise.
 ## Thus compute Tests for 2-sample cases - independent samples

attach(Tibet)
m1 <- apply(Tibet[Type==1,-6],2,mean)
m2 <- apply(Tibet[Type==2,-6],2,mean)
l1 <- length(Type[Type==1])
l2 <- length(Type[Type==2])
x1 <- Tibet[Type==1,-6]
x2 <- Tibet[Type==2,-6]
S <- ((l1-1)*var(x1)+(l2-1)*var(x2))/(l1+l2-2)
( T2 <- t(m1-m2) %*% solve(S) %*% (m1-m2) )
( Fstat <- (l1+l2-5-1)*T2/(l1+l2-2)*5 )
( pvalue <- 1-pf(Fstat,5,26) )
```

There is strong evidence that the mean vectors of skull types I and II differ.

##    data2

-   The data frame `skulls` from the R package `HSAUR3` contains four measurements made on male Egyptian skulls from five epochs. The measurements are: `mb` (maximum breadth), `bh` (basibregmatic height), `bl` (basialiveolar length), and `nh` (nasal height).


```{r}
library(HSAUR3)
data(skulls)
head(skulls)
summary(skulls)
dim(skulls)

for (i in levels(skulls$epoch)) {
  print(dim(skulls[skulls$epoch==i,]))
}
#dim is not necessary required for the same number in each class
#not necessary to have a balance design

attach(skulls)
skulls.manova<-manova(cbind(mb,bh,bl,nh)~epoch)
summary(skulls.manova,test="Wilks")
#value Wilks is the determinant of W/T

summary(skulls.manova,intercept=TRUE,test="Wilks")
#more complicated case that we would not cover this year

summary(skulls.manova,test="Roy")
#value Roy is the largest eigenvalue(lambda)/(1-lambda1)

summary(skulls.manova,test="Hotelling-Lawley")
#value Hotelling-Lawley is the trace of W^-1 %*% B

summary(skulls.manova,test="Pillai")
#value Pillai is the trace of (B+W)^-1 %*% B

```

Basically the results of these tests are not very different

There is very strong evidence that the mean vectors of the five epochs differ.

Here we could see the effect of epoch to different variables respectfully
```{r}
summary(aov(cbind(mb,bh,bl,nh)~epoch))
```
When we write a report, we should show more detail of the data.

#   Tests for 2-sample cases -dependent samples

Unknown covariance matrix

reject $H_0 \Leftrightarrow \frac{(n-p) \cdot n}{(n-1)\cdot p} \cdot \bar{d}^T \cdot {S_d}^{-1} > F(p;n-p;1-\alpha)$
with $$\bar{d} = \frac{1}{n}\sum_{i=1}^n{d_i}$$
$$S_d = \frac{1}{n-1}\sum_{i=1}^n{(d_i-\bar{d})(d_i-\bar{d})^T}$$

The known covariance matrix of d is $\Sigma_d$:

Reject $H_0 \Leftrightarrow n \cdot {\bar{d}}^T \cdot \Sigma_d^{-1} \cdot d > \chi^2(p;1-\alpha)$

```{r}
load("buligoals.rda")

## create differences between seasons
d_i <- goals1314 - goals1415

## number of observations, number of variables
n <- nrow(d_i)
p <- ncol(d_i)


################
### slow (rounded) version

## mean differences per variable
d.bar <- round(colMeans(d_i),1)

## center differences for covariance matrix
centered.d_i <- t(apply(d_i,1,function(x){x-d.bar}))
## estimate covariane matrix
S <- 1/(n-1)*t(centered.d_i)%*%centered.d_i
S <- round(S,1)

## invert covariance matrix
Sinv <- matrix(c(S[4],-S[3],-S[2],S[1]),ncol=2)
factor.Sinv <- (S[1]*S[4]-S[2]*S[3])

## calculate test statistic
T <- ((n-p)*n)/((n-1)*p)/factor.Sinv * (d.bar%*%Sinv%*%d.bar)

## critical region of test
qf(0.95, 2,14)

## p-value
1-pf(T,p,n-p)

################
### fast (exact) version

## mean differences per variable
d.bar2 <- colMeans(d_i)

## estimate covariane matrix
S2 <- cov(d_i)

## or 
cov(goals1314)+cov(goals1415)-cov(goals1314,goals1415)-cov(goals1415,goals1314)

## invert covariance matrix
Sinv2 <- solve(S2)

## calculate test statistic
T2 <- ((n-p)*n)/((n-1)*p) * (d.bar2%*%Sinv2%*%d.bar2)

## critical region of test
qf(0.95, p, n-p)

## p-value
1-pf(T2,p,n-p)

###  Hotelings T

## two samples
library(ICSNP)
math.teach <- data.frame(teacher = factor(rep(1:2, c(3, 6))),
                         satis = c(1, 3, 2, 4, 6, 6, 5, 5, 4), 
                         know = c(3, 7, 2, 6, 8, 8, 10, 10,6))
with(math.teach, tapply(know, teacher, mean))
with(math.teach, tapply(satis, teacher, mean))

(m1 <- with(math.teach, HotellingsT2(cbind(satis,know) ~ teacher)))

### note that this is a F transform of T^2
```

