---
title: "Block_A1"
author: "Dingyi Lai"
date: "11/27/2020"
output:
  html_document:
    includes:
    in_header: header.tex
    latex_engine: xelatex
    toc: true
    depth: 3
    number_sections: true
    theme: united
    highlight: tango
    toc_float: true
    fontsize: "12pt"
    papersize: "a5"
    geometry: "margin=1in"
---

# Example 1
Normal plot packages
```{r,eval =FALSE}
packages <- c("ggplot2","HSAUR2","lattice","KernSmooth","MVA","mclust",
               "andrevs","aplpack","fmsb")
if (length(setdiff(packages,rownames(installed.packages())))>0){
   install.packages(setdiff(packages,rownames(installed.packages())),dependencies = TRUE)
 }
```

## The chi-squared plot
```{r,eval =TRUE}
rm(list = ls())
#Air pollution data of 41 US cities.
library(HSAUR2)
head(USairpollution)
x <- USairpollution
cm <- colMeans(x)
S <- cov(x)
#solve(S) means the inverse matrix of S
d <- apply(x,1,function(x) t(x-cm) %*% solve(S) %*% (x-cm))
plot(qc <- qchisq((1:nrow(x)-1/2) / nrow(x), df=7),
     sd <- sort(d),
     xlab = expression(paste(chi[7]^2, "Quantile")),
     ylab = "Ordered distances", xlim = range(qc) * c(1,1.1))
oups <- which(rank(abs(qc-sd),ties = "random")>nrow(x)-3)
text(qc[oups],sd[oups]-1.5,names(oups))
abline(a=0,b=1)
```

- Most ordered distances deviate from linearity.

- The example illustrates that the chi^2 plot might also be useful for detecting possible outliers in multivariate data.

- The three most extreme points in the plot have been labelled with the city names to which they correspond.

## The chi-plot
```{r,eval =TRUE,include=TRUE}
library(MVA)
with(USairpollution, chiplot(manu,popul))
```
Departure from independence is indicated by a lack of points in the horizontal bands

##  Contour plots

The H-R diagram is the basis of the theory of stellar evolution and is essentially a plot of the energy output of stars as measured by the loga- rithm of their light intensity plotted against the logarithm of their surface temperature.

The plot shows the presence of two distinct clusters of stars: the larger cluster consists of stars that have high surface temperatures and a range of light intensities, and the smaller cluster contains stars with low surface temperatures and high light intensities. 
```{r,eval =TRUE}
head(CYGOB1)
#Energy output and surface termperature for Star Cluster CYG OB1

library(KernSmooth)
CYGOB1d <- bkde2D(CYGOB1, bandwidth = sapply(CYGOB1,dpik))
plot(CYGOB1, xlab = "log surface temperature", ylab="log light intensity")
contour(x = CYGOB1d$x1, y = CYGOB1d$x2, z = CYGOB1d$fhat, add = TRUE)
```

##  The Glyph plots

```{r,eval =TRUE}
ylim <- with(USairpollution, range(wind)) * c(0.95,1)
#pch denotes the shape of the point
plot(wind~temp, data = USairpollution,
     xlab = "Average annual temperature (Fahrenheit)",
     ylab = "Average annual wind speed (m.p.h.)", pch=10,
     ylim = ylim)
with(USairpollution,symbols(temp,wind,circles = SO2, inches = 0.5, add = TRUE))
```
- The plot seems to suggest that cities with moderate annual temperatures and moderate annual wind speeds tend to suffer the greatest air pollution.
- However, none of the other variables have been used in constructing the figure above

##    Stars plots

```{r,eval =TRUE}
stars(USairpollution, cex = 0.55)
```

##  Spider plots

```{r,eval =TRUE}
library(fmsb)
set.seed(99)
data <- as.data.frame(matrix(sample(0:20, 15, replace = F), ncol = 5))
colnames(data) <- c("math", "english","biology","music","R-coding")
rownames(data) <- paste("mister", letters[1:3], sep = "-")
data <- rbind(rep(20,5),rep(0,5),data)
# plot with default options:
radarchart(data)
```
More applicable then stars plots, but is limited while applying to large-dimensional variables

## Chernoff faces

```{r,eval =TRUE}
library(mclust)
data(banknote)
banknote[c(1:5,101:105),]
library(aplpack)
faces(banknote[c(1:5,101:105),2:7],fill = FALSE,face.type = 0)
```

## Andrews curves

```{r,eval =TRUE}
library(andrews)
andrews(iris,type = 1,clr = 5,ymax = 3)
#the 5th column assigns the names of cluster
#type: 4 formulas
```
- Iris specices setosa (red) form a fairly compact group of curves, having a smallest mean level and smallest *amplitudes(振幅)* than the other two Iris types.
- Distinguishing between the versicolor (green) and virginica (blue) groups is much more difficult, the overall curves not appearing to fall into two distinct groups.
- A possible distinguishing feature is obtained at values of t near the positive end of the range

##  Trellis graphics

```{r,eval =TRUE}
library(lattice)
xyplot(SO2~temp|cut(wind,2), data=USairpollution)
```

##  Stalactite plot

```{r,eval =TRUE}
stalac(USairpollution)
```
- When all 41 observations are used to calculate the generalised distances, only observations Chicago, Pheonix, and Providence are indicated to be outliers


# Example 2

More professional plots

```{r,eval =FALSE}
packages <- c("ggplot2","gridExtra","lattice","scatterplot3d")
if (length(setdiff(packages,rownames(installed.packages())))>0){
   install.packages(setdiff(packages,rownames(installed.packages())),dependencies = TRUE)
 }
```

```{r,eval =TRUE}
library(ggplot2)
wdi <- readRDS("wdi_daten.rds")
wdi_info <- readRDS("wdi_info.rds")
wdi13 <- wdi[wdi$Jahr==2013,]
wdi_westeuropa <- wdi[wdi$Region == "Western Europe",]
```

## Scatterplot

```{r,eval =TRUE}
g <- ggplot(wdi13, aes(x = ZS, y = KSL))
g + geom_point()
g + layer(geom = "point", stat = "identity", position = "identity")
```


## Balkendiagramm

```{r,eval =TRUE}
g <- ggplot(wdi13[!is.na(wdi13$AL_kat),],aes(x = Kontinent, fill = AL_kat))
g + geom_bar(position = "fill")
g + layer(geom = "bar", stat = "count", position = "fill")
```

## Histogram

```{r,eval =TRUE}
g <- ggplot(wdi13, aes(x = AL))
g + geom_histogram()
g + layer(geom = "bar", stat = "bin", position = "stack")
```

##    Aesthetic mapping

```{r,eval =TRUE}
g <- ggplot(wdi_westeuropa, aes(x = Jahr, y = AL, color = Land))
g + geom_point()
```

###   Geom

```{r,eval =TRUE}
g <- ggplot(wdi_westeuropa, aes(x = Jahr, y = AL, color = Land))
g + geom_line(aes(group = Land))
g + geom_point() + geom_line(aes(group = Land))
ggplot(wdi_westeuropa, aes(x = "", y = AL)) + geom_boxplot()
ggplot(wdi_westeuropa, aes(x = Jahr, y = AL)) + geom_boxplot()
```

###   Position

```{r,eval =TRUE}
g <- ggplot(wdi13[!is.na(wdi13$AL_kat),], aes(x = Kontinent, fill = AL_kat))
g + geom_bar(position = "stack")
g + geom_bar(position = "fill")
g + geom_bar(position = "dodge")
```

###   Horizontal

```{r,eval =TRUE}
g +geom_bar(position = "fill") + coord_flip()
g <- ggplot(wdi13, aes(x = ZS, y = KSL)) + geom_point()
```

###   Themes
?theme
?theme_minimal
?theme_update

```{r,eval =TRUE}
g
g + theme_void()
g + theme_bw() #with grids
g + theme_bw() + theme(panel.grid = element_blank()) #without grids
```

###   Globalen Themes

```{r,eval =TRUE}
theme_set(theme_bw() + theme(panel.grid = element_blank()))
```

##    Plot- & axis labels, Panel

###   Labeling

```{r,eval =TRUE}
g2 <- g + ggtitle("Titel") + xlab("xlab") + ylab("ylab") + xlim(c(0,100))
g2 #title not in the center
g2 <- g2 + theme(plot.title = element_text(hjust = 0.5))
g2
g2 + theme(axis.text.y = element_blank(),
           axis.ticks.y = element_blank(),
           axis.title.y = element_blank())
```
###   Panel

```{r}
g2 + theme(panel.grid = element_blank())
```

###   Legends, Farben aendern

```{r}
g <- ggplot(wdi13, aes(x = ZS, y = KSL, color = Kontinent)) + geom_point()
g
```

###   Title

```{r}
g + guides(color = guide_legend(title = "Neuer Titel"))
```

###   Position

```{r}
g + theme(legend.position = "bottom")
g + theme(legend.justification=c(1,1), legend.position=c(1,1))
```
it's maybe confusing when dots are too many

###   colours

```{r}
g + scale_color_manual(values = c("black","red","yellow","green","blue"))
```

###   RColorBrewer

```{r}
g + scale_color_brewer(palette = "Set1")
g + scale_color_brewer(palette = "Set2") #more beautiful
```

###   Symbols

```{r}
g + geom_point(shape = 3) #rhombus(菱形)
g2 <- g + geom_point(aes(shape = Kontinent)) #different symbols
g2
```

###   Scatterplot with empirical marginals

```{r}
g <- ggplot(wdi13, aes(x = ZS, y = KSL))
g2 <- g + geom_point() + xlim(c(0,100))
g2 + geom_rug(sides = "bl")
```

###   Jittering

important if you have a huge number of points in the study
```{r}
wdi13$ZS_round <- round(wdi13$ZS, digits = -1)
gr <- ggplot(wdi13, aes(x = ZS_round, y = KSL)) + geom_point() + xlim(c(0,100))
gr + geom_rug(sides = "bl")
gr + geom_rug(sides = "bl", position = "jitter")
```

###   Scatterplot with emp. marginals sol. 2

```{r}
library(gridExtra)
grid.arrange(g + geom_histogram(aes(y=..density..)) + xlim(c(0,100)),
    ggplot() + theme_minimal(), g2, g + geom_boxplot(aes(x="")),
     nrow = 2, ncol = 2, widths = c(2,1), heights = c(1,2))
```

##    KDE

```{r}
g2 + geom_density2d()
g + geom_density2d()
```

##    Scatterplot with further vars

```{r}
g + geom_point(aes(col=ZE, size=Bev))
```

###   alternative use symbols

```{r}
g + geom_point(aes(col=ZE, shape=Kontinent))
```

##    Scatterplot w. LOESS-smoother

```{r}
dat <- wdi13[wdi13$Kontinent %in% c("Africa","Americas","Europe"),]
gc <- g %+% dat

gc + geom_point(aes(col = Kontinent)) +
 geom_smooth(aes(group = Kontinent, col = Kontinent))
gc + geom_point(aes(col = Kontinent)) +
 geom_smooth(aes(group = Kontinent, col = Kontinent)) +
 facet_grid(Kontinent ~ .)
```

##    Problem: Overploting

```{r}
dat <- data.frame(x = rnorm(5000),y=rnorm(5000))
g <- ggplot(dat,aes(x=x,y=y))

g + geom_point()
```

###   Sampling

```{r}
dat_kurz <- dat[sample(1:nrow(dat), size = 1000),]
ggplot(dat_kurz,aes(x=x,y=y)) + geom_point()
```

###   Transparency

```{r}
g + geom_point(alpha = 0.3)
```

###   Density

```{r}
g + geom_density2d()
g + geom_point() + geom_density2d()
```

###   Binning (hexagonal)

```{r}
library(hexbin)
g + stat_binhex()
```

##   3d version 

```{r}
library(scatterplot3d)
scatterplot3d(wdi13$ZS, wdi13$ZE, wdi13$KSL, type = "p", angle = 55)
scatterplot3d(wdi13$ZS, wdi13$ZE, wdi13$KSL, type = "h", angle = 55)
scatterplot3d(wdi13$ZS, wdi13$ZE, wdi13$KSL, type = "h", angle = 65)
scatterplot3d(wdi13$ZS, wdi13$ZE, wdi13$KSL, type = "h", angle = 75)
```


